Tuning for Software Analytics: is it Really Necessary?

### 2016-04-03 tuningpredictor.tex
Updates from ShareLaTeX

\begin{document}
\begin{frontmatter}
\title{ Tuning for Software Analytics: is it Really Necessary?}
\author{Wei Fu\corref{cor1}}
\ead{wfu@ncsu.edu}
\author{Tim Menzies\corref{cor1}}
\ead{tim.menzies@gmail.com}
\author{Xipeng Shen}
\ead{ xshen5@ncsu.edu}
\cortext[cor1]{Corresponding author: Tel.:+1 304-376-2859 (Tim Menzies)}
\address{Department of Computer Science, North Carolina State University, Raleigh, NC, USA}

@@

% \thispagestyle{plain}
% \pagestyle{plain}
\begin{abstract}
\textbf{Context:} Data miners have been widely used to generate defect predictors from static code measures. Static code defect predictors perform well compared to manual methods, and they are easy to use and useful to use.  \\
\textbf{Objective:} One of the ``black arts'' of data mining is setting the tuning that control the miner. We offer a simple, automatic, and very effective  method for finding those tunings.\\
\textbf{Method:} For each experiment with different data sets, we run differential evolution as an optimizer to explore the tuning space as the first step and  fit the data miners with the best tunings afterwards.\\
\textbf{Results:} Contrary to our prior expectations, we found these tunings
 was remarkably simple: it only required tens, not thousands, of attempts
 to obtain very good results.For example, when  learning software defect predictors, this  method can quickly
 find tunings  that  alter detection  precision
 from 0\% to 60\%. \\
\textbf{Conclusion:} Given that (1)~the improvements are so large, and (2)~the tuning is so simple,
these results prompt for a change to standard methods in software analytics.
At least for defect prediction, 
it is no longer enough to just run a data miner and present the result
{\em without}  conducting a tuning optimization study.
The implication for other kinds of  analytics is now  an open and pressing issue.

@@

The results of this paper show that the above points are
false,  at least for
defect prediction from  code attributes:
\be
\item
Tuning  defect predictors is {\em remarkably simple};
\item
And can {\em dramatically improve the performance}. 
\ee
Those results were found by   exploring six research questions:
\bi
\item RQ1: {\em Does   tuning    improve the performance scores of a predictor?} We will show below
 examples of truly dramatic improvement:
 usually by 5 to 20\% and often by much more (in one extreme case,  precision improved from 0\% to 60\%).
\item RQ2: {\em Does tuning change conclusions on what learners are better than others?} 
Recent SE papers~\cite{lessmann2008benchmarking,hall11} claim that some learners are better than others. 
Some of those conclusions are completely changed by tuning. 
\item RQ3: {\em Does tuning change conclusions about what factors are most important in software engineering?} Numerous recent SE papers (e.g.~\cite{bell2013limited,rahman2013how,me02k,Moser:2008,zimmermann2007predicting,%
herzig2013predicting}) use data miners to conclude that {\em this}
is more important than {\em that} for reducing software project defects.
Given the  tuning results of this paper, we show that such conclusions need to be revisited.
\item  RQ4: {\em Is tuning easy?} We show that one of the simpler multi-objective optimizers
(differential evolution~\cite{storn1997differential}) works very well for tuning defect predictors. 
\item RQ5: {\em Is tuning impractically slow?} We achieved dramatic improvements in the performance scores
of our data miners in less than 100 evaluations (!); i.e., very
quickly.
\item RQ6: {\em Should data miners be used ``off-the-shelf'' with their default tunings?} 
For defect prediction from static code measures, our answer is an emphatic ``no'' (and
the implication for other kinds of  analytics is now an open and urgent question).
\ei
Based on our answers to these  questions,  we strongly advise that:
\bi
\item
Data miners should not be used ``off-the-shelf'' with default tunings.
\item
Any future paper on defect prediction should include a 
tuning study. Here, we have found  an algorithm called differential
evolution to be a useful method for conducting such
tunings.
\item
Tuning needs to be repeated
whenever data or goals are changed.
Fortunately, the cost of finding good tunings is not excessive since, at least for
static code defect predictors, tuning is easy and fast.
\ei

@@

\subsection{Tuning: Important and Ignored}
Having made the business case for tuning with respect to the business context,
we now show that tuning is an under-explored area in   SE.

@@

In the following, we present tuning as a multi-objective optimization problem.
With such optimizations,  it is not always possible to optimize for every goal.
For example, the following text does not
show results for tuning on recall
or false alarms since optimizing {\em only} for those goals can lead
to some undesirable side effects:
\bi
\item
{\em Recall} reports the percentage of  predictions that are actual examples of  what we are looking for.
When we tune for {\em recall}, we can achieve near
100\% recall-- but the cost of a near 100\% false alarms.
\item
{\em False alarms} is the percentage of other examples that are reported  (by the learner)
to be part of the targeted examples.
When we tune for {\em false alarms}, we 
can achieve near zero percent false alarm rates by effectively turning off
the detector (so the  recall falls to nearly zero).
\ei
Accordingly,  this paper  explores performance measures that comment on all 
target classes: see the  precision and ``F'' measures discussed below: see {\em Optimization Goals}.
That said, we are sometimes asked what good is a learner if it optimizers for (say) precision
at the expense of (say) recall.

Our reply is that software engineering is a very diverse enterprise
and that different kinds of development need to optimize for different goals
(which may not necessarily be ``optimize for recall''):
\bi
\item
Anda, Sjoberg and Mockus are concerned with {\em reproducibility}  and so
assess their models using the the ``coefficient of variation'' $C\frac{stddev}{mean}$) ~\cite{anda09}.
\item
Arisholm~\&~Briand~\cite{arisholm06},  Ostrand \& Weyeuker~\cite{ostrand04} and Rahman et al.~\cite{rahman12} are concerned with reducing the work load associated with someone
else reading a learned model, then applying it. Hence, they assess their models using  {\em reward}; i.e.   the fewest lines of code
  containing the most bugs.
\item
Yin et al. are concerned about
 {\em incorrect bug fixes}; i.e. those that require subsequent work in order to complete the bug fix.
These bugs occur  when (say) developers try to fix parts of the code
where they have very little experience~\cite{yin11}.  Hence, they assess a learned
model using a measure that selects for  the most number of bugs in regions that {\em the most programmers have worked with before}.
\item
For safety critical applications, high false alarm rates are  acceptable if the cost
of overlooking  critical issues outweighs the inconvenience of   inspecting a few more
modules. 
\item
When rushing a product to market,  there is a business case to 
avoid the extra rework associated with false alarms.  In that business context, 
managers might be willing to lower the recall somewhat in order to minimize the false alarms.
\item
When the second author worked with contractors at  NASA's software independent verification
and validation facility, he found  new contractors  
only reported issues that were most certainly important defects; i.e. they minimized
  false alarms even if that damaged their precision (since, they felt, 
it was better to silent than wrong). Later on, once
those contractors had acquired a reputation of being insightful members of the team,
they improved their precision scores (even if it means some more false alarms).
\ei
Accordingly, this paper does not assume that (e.g.) minimizing false alarms is 
more important than maximizing   precision. Such a determination 
depends on   business conditions.
Rather, what we can  show  examples where  changing  optimization goals can also change 
the conclusions made from that learner on that data. More generally, we caution that it is 
important not to overstate  empirical results from  analytics.
Rather, those results need to be expressed {\em along with} the context within which they are
relevant (and by ``context'', we mean the optimization goal).

@@

This section is our standard introduction to defect prediction~\cite{me15:book1},
plus    some new results from Rahman et al.~\cite{rahman14:icse}.

@@

\renewcommand{\baselinestretch}{0.8}
\scriptsize
\centering
  \begin{tabular}{p{0.75cm}p{0.75cm}p{0.75cm}p{0.75cm}p{0.75cm}p{0.75cm}p{0.75cm}p{0.75cm}p{0.75cm}p{0.75cm}}\hline
  Dataset &antV0&antV1&antV2&camelV0&camelV1&ivy&jeditV0&jeditV1&jeditV2
\\\hline
  training &20/125 &40/178 &32/293 &13/339 &216/608 &63/111 &90/272 &75/306 &79/312
\\  tuning  &40/178 &32/293 &92/351 &216/608 &145/872 &16/241 &75/306 &79/312 &48/367
\\  testing &32/293 &92/351 &166/745 &145/872 &188/965 &40/352 &79/312 &48/367 &11/492
\\ \hline
  Dataset &log4j&lucene&poiV0&poiV1&synapse&velocity&xercesV0&xercesV1
\\\hline
  training &34/135 &91/195 &141/237 &37/314 &16/157 &147/196 &77/162 &71/440
\\  tuning  &37/109 &144/247 &37/314 &248/385 &60/222 &142/214 &71/440 &69/453
\\  testing &189/205 &203/340 &248/385 &281/442 &86/256 &78/229 &69/453 &437/588
\\  \end{tabular}

@@

\fig{deltas} says  that the answer to RQ1 is ``yes''-- tuning  has a positive effect on performance scores. This figure sorts
 deltas in the precision and the F-measure    between tuned and untuned learners. Our reading of this
figure is that, overall, tuning rarely makes performance   worse and usually can make it much better.

>>>>

\begin{document}
\begin{frontmatter}
\title{ Tuning for Software Analytics: is it Really Necessary?}
\author{Wei Fu\corref{cor1}}
\ead{wfu@ncsu.edu}
\author{Tim Menzies\corref{cor1}}
\ead{tim.menzies@gmail.com}
\author{Xipeng Shen}
\ead{ xshen5@ncsu.edu}
\cortext[cor1]{Corresponding author: Tel:19195345251(Wei)}
\address{Department of Computer Science, North Carolina State University, Raleigh, NC, USA}

@@

% \thispagestyle{plain}
% \pagestyle{plain}
\begin{abstract}
\textbf{Context:} Data miners have been widely used in
software engineering to, say,  generate defect predictors from static code measures. Such static code defect predictors perform well compared to manual methods, and they are easy to use and useful to use.  But 
one of the ``black arts'' of data mining is setting the tunings that control the miner. \\
\textbf{Objective:} We seek  simple, automatic, and very effective  method for finding those tunings.\\
\textbf{Method:} For each experiment with different data sets (from open source JAVA systems), we ran differential evolution as an optimizer to explore the tuning space
(as a first step) then tested the tunings
using hold-out data.\\
\textbf{Results:} Contrary to our prior expectations, we found these tunings
 were remarkably simple: it only required tens, not thousands, of attempts
 to obtain very good results. For example, when  learning software defect predictors, this  method can quickly
 find tunings  that  alter detection  precision
 from 0\% to 60\%. \\
\textbf{Conclusion:} Since the (1)~the improvements are so large, and (2)~the tuning is so simple, we need 
to change  standard methods in software analytics.
At least for defect prediction, 
it is no longer enough to just run a data miner and present the result
{\em without}  conducting a tuning optimization study.
The implication for other kinds of  analytics is now  an open and pressing issue.

@@

The results of this paper show that the above points are
false since, at least for
defect prediction from  code attributes:
\be
\item
Tuning  defect predictors is {\em remarkably simple};
\item
And can {\em dramatically improve the performance}. 
\ee
Those results were found by   exploring six research questions:
\bi
\item RQ1: {\em Does   tuning    improve the performance scores of a predictor?} We will show below
 examples of truly dramatic improvement:
 usually by 5 to 20\% and often by much more (in one extreme case,  precision improved from 0\% to 60\%).
\item RQ2: {\em Does tuning change conclusions on what learners are better than others?} 
Recent SE papers~\cite{lessmann2008benchmarking,hall11} claim that some learners are better than others. 
Some of those conclusions are completely changed by tuning. 
\item RQ3: {\em Does tuning change conclusions about what factors are most important in software engineering?} Numerous recent SE papers (e.g.~\cite{bell2013limited,rahman2013how,me02k,Moser:2008,zimmermann2007predicting,%
herzig2013predicting}) use data miners to conclude that {\em this}
is more important than {\em that} for reducing software project defects.
Given the  tuning results of this paper, we show that such conclusions need to be revisited.
\item  RQ4: {\em Is tuning easy?} We show that one of the simpler multi-objective optimizers
(differential evolution~\cite{storn1997differential}) works very well for tuning defect predictors. 
\item RQ5: {\em Is tuning impractically slow?} We achieved dramatic improvements in the performance scores
of our data miners in less than 100 evaluations (!); i.e., very
quickly.
\item RQ6: {\em Should data miners be used ``off-the-shelf'' with their default tunings?} 
For defect prediction from static code measures, our answer is an emphatic ``no'' (and
the implication for other kinds of  analytics is now an open and urgent question).
\ei
Based on our answers to these  questions,  we strongly advise that:
\bi
\item
Data miners should not be used ``off-the-shelf'' with default tunings.
\item
Any future paper on defect prediction should include a 
tuning study. Here, we have found  an algorithm called differential
evolution to be a useful method for conducting such
tunings.
\item
Tuning needs to be repeated
whenever data or goals are changed.
Fortunately, the cost of finding good tunings is not excessive since, at least for
static code defect predictors, tuning is easy and fast.
\ei

@@

\subsection{Tuning: Important and Ignored}
This section argues that   tuning is an under-explored software analytics-- particularly in the apparently  well-explored field of defect prediction.

@@

 Having made the case that tuning needs to be explored more,
 but before we get into the technical details of this
 paper, this section discusses some
 general matters about setting goals during tuning
 experiments.

 This paper characterizes tuning as an optimization problem (how to change the settings on the learner
 in order to best improve the output).
With such optimizations,  it is not always possible to optimize for all goals at the same time.
For example, the following text does not
show results for tuning on recall
or false alarms since optimizing {\em only} for those goals can lead
to some undesirable side effects:
\bi
\item
{\em Recall} reports the percentage of  predictions that are actual examples of  what we are looking for.
When we tune for {\em recall}, we can achieve near
100\% recall-- but the cost of a near 100\% false alarms.
\item
{\em False alarms} is the percentage of other examples that are reported  (by the learner)
to be part of the targeted examples.
When we tune for {\em false alarms}, we 
can achieve near zero percent false alarm rates by effectively turning off
the detector (so the  recall falls to nearly zero).
\ei
Accordingly,  this paper  explores performance measures that comment on all 
target classes: see the  precision and ``F'' measures discussed below: see {\em Optimization Goals}.
That said, we are sometimes asked what good is a learner if it optimizers for (say) precision
at the expense of (say) recall.

Our reply is that software engineering is a very diverse enterprise
and that different kinds of development need to optimize for different goals
(which may not necessarily be ``optimize for recall''):
\bi
\item
Anda, Sjoberg and Mockus are concerned with {\em reproducibility}  and so
assess their models using the the ``coefficient of variation'' $C\frac{stddev}{mean}$) ~\cite{anda09}.
\item
Arisholm~\&~Briand~\cite{arisholm06},  Ostrand \& Weyeuker~\cite{ostrand04} and Rahman et al.~\cite{rahman12} are concerned with reducing the work load associated with someone
else reading a learned model, then applying it. Hence, they assess their models using  {\em reward}; i.e.   the fewest lines of code
  containing the most bugs.
\item
Yin et al. are concerned about
 {\em incorrect bug fixes}; i.e. those that require subsequent work in order to complete the bug fix.
These bugs occur  when (say) developers try to fix parts of the code
where they have very little experience~\cite{yin11}.  Hence, they assess a learned
model using a measure that selects for  the most number of bugs in regions that {\em the most programmers have worked with before}.
\item
For safety critical applications, high false alarm rates are  acceptable if the cost
of overlooking  critical issues outweighs the inconvenience of   inspecting a few more
modules. 
\item
When rushing a product to market,  there is a business case to 
avoid the extra rework associated with false alarms.  In that business context, 
managers might be willing to lower the recall somewhat in order to minimize the false alarms.
\item
When the second author worked with contractors at  NASA's software independent verification
and validation facility, he found  new contractors  
only reported issues that were most certainly important defects; i.e. they minimized
  false alarms even if that damaged their precision (since, they felt, 
it was better to silent than wrong). Later on, once
those contractors had acquired a reputation of being insightful members of the team,
they improved their precision scores (even if it means some more false alarms).
\ei
Accordingly, this paper does not assume that (e.g.) minimizing false alarms is 
more important than maximizing   precision or recall. Such a determination 
depends on   business conditions.

Rather, what we can  show  examples where  changing  optimization goals can also change 
the conclusions made from that learner on that data. More generally, we caution that it is 
important not to overstate  empirical results from  analytics.
Those results need to be expressed {\em along with} the context within which they are
relevant (and by ``context'', we mean the optimization goal).

@@

This section discusses defect prediction,
which is the particular
task explored by our optimizers.
Note that this section repeats much of 
our standard introduction to defect prediction~\cite{me15:book1},
as well as presenting    some new results from Rahman et al.~\cite{rahman14:icse}.

@@

\renewcommand{\baselinestretch}{0.8}
\scriptsize
\centering
%   \begin{tabular}{p{0.75cm}p{0.75cm}p{0.75cm}p{0.75cm}p{0.75cm}p{0.75cm}p{0.75cm}p{0.75cm}p{0.75cm}p{0.75cm}}\hline
  \begin{tabular}{c c c c c c c c c c } \hline
  Dataset &antV0&antV1&antV2&camelV0&camelV1&ivy&jeditV0&jeditV1&jeditV2
\\\hline
  training &20/125 &40/178 &32/293 &13/339 &216/608 &63/111 &90/272 &75/306 &79/312
\\  tuning  &40/178 &32/293 &92/351 &216/608 &145/872 &16/241 &75/306 &79/312 &48/367
\\  testing &32/293 &92/351 &166/745 &145/872 &188/965 &40/352 &79/312 &48/367 &11/492
\\ \hline
  Dataset &log4j&lucene&poiV0&poiV1&synapse&velocity&xercesV0&xercesV1
\\\hline
  training &34/135 &91/195 &141/237 &37/314 &16/157 &147/196 &77/162 &71/440
\\  tuning  &37/109 &144/247 &37/314 &248/385 &60/222 &142/214 &71/440 &69/453
\\  testing &189/205 &203/340 &248/385 &281/442 &86/256 &78/229 &69/453 &437/588
\\  \end{tabular}

@@

\fig{deltas} says  that the answer to RQ1 is ``yes''-- tuning  has a positive effect on performance scores. This figure sorts
 deltas in the precision and the F-measure    between tuned and untuned learners. Our reading of this
figure is that, overall, tuning rarely makes performance   worse and often can make it much better.
